---
title: "Titanic_kaggle"
author: "Jeff Kropelnicki"
date: "11/16/2016"
output: html_document
  
---

```{r message=FALSE, warning=FALSE}

library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
```


Load data, bind the two together and show data. 
```{r, message=FALSE, warning=FALSE}
train <- read.csv("/Users/jeffkropelnicki/Documents/kaggle/Exploring the Titanic Dataset/train.csv", stringsAsFactors = F)
test <- read.csv("/Users/jeffkropelnicki/Documents/kaggle/Exploring the Titanic Dataset/test.csv", stringsAsFactors = F)
full <- bind_rows(train, test) #binds train and test 
str(full)
```   
    
|**Variable Name** |**Description**                 
| :------------- |------------------------
| Survived      | [0] = Died [1] = Survived 
| Pclass        | Passenger's class           
| Name          | Passenger's name         
| Sex           | Passenger"s sex          
| Age           | Passenger's name
| SibSp         | Number of siblings/spouses aboard
| Parch         | Number of parents/children aboard
| Ticket        | Ticket number 
| Fare          | Fare
| Cabin         | Cabin
| Embarked      | Port of embarkation   
   
Looking at the names we can see that some have Mr, Mrs, Master, Miss. A quest we could ask is does this say anything about the passengers class? If we pull out the title this will give us a new varible. We could also use sir name to look for family so I will pull that too.   


```{r}
# Get title from name 
full$Title <- gsub('(.*, )|(\\..*)', '', full$Name)
# Show title count by sex   
table(full$Sex, full$Title)   
# Combinding tiltle with low counts. 
rare_title <- c("Capt", "Col", "Don", "Dona", "Dr", "Jonkheer", "Lady", "Major", "Rev", "Sir", "the Countess")   
# Reassign mlle, ms, and mme accordingly   
full$Title[full$Title == 'Mlle']  <- "Miss"
full$Title[full$Title == "ms"] <- "Miss"
full$Title[full$Title == "mme"] <- "Mrs"
full$Title[full$Title %in% rare_title] <- "Rare Title"
# looking at the changes made to the table above
table(full$Sex, full$Title) 
 # now I want to look at sir names   
full$Surname <- sapply(full$Name,  
                      function(x) strsplit(x, split = '[,.]')[[1]][1])
```      
```{r results='asis'}
cat(paste('There are', nlevels(factor(full$Surname)), 'unique surnames.'))
```   
Now I want to make a new family varible.   
```{r}
# Create a family size variable including the passenger themselves
Temp_family_size <- full$SibSp + full$Parch + 1
# Create a family variable 
full$Family <- paste(full$Surname, full$Fsize, sep='_')

```















